The binary liquid model for gravity-driven microchannel simulations

### 2010-10-22 Text/paper.tex
Added grid results to the text.

The example what we want to simulate is the capillary number
$0.05$. As it was pointed out the film thickness in that case is $6$ percent.
Note that in the case of half-way bounce-back walls which are used in the
simulations one needs to calculate the film thickness as: 
\begin{equation}
\delta=(\phi_0-0.5)/(N_y-2),
\end{equation}
where $\phi_0$ is the grid coordinate where phase field equals to $0$, $N_y-2$
is related to the effective channel height. Note that it's a big
simplifications to impose the boundary in the middle beetween the bounce-back
node and the fluid node. In the case of $\delta=0.06$ the initial grid number
is $N_y=102$, which gives the horizontal grid number as $N_x-1=15(N_y-2)=1500$.
The bubble is initialized as rectangular box with coordinates
$y=7\dots N_y-8$, $x=\frac{N_x}{3}\dots \frac{2 N_x}{3}$ and phase
$\phi_{bubble}=-1$. All other nodes are initialized with the phase field
$\phi=1$. The force gradient can be estimated through the Poiseuille
profile formula, Eq.\ref{poiseuille:velocity:center}, and it equals
$1.508\,10^{-6}$.

After choosing the reference parameters, the grid refinement needs to keep the
same the same macroscopic parameters as capillary number and viscosities ratio.
Therefore the simulations were performed with grid
numbers $N_y=102,127,152,177,202,227$, corresponding to the
effective channels widths as $H_{eff}=100,125,150,175,200,225$,
$N_x=1501,1876,2251,2626,3001,3376=15
H_{eff}+1$. It is easy to check
that keeping all liquid viscosities the same yields the following quantity to
be a
constant while doing the grid refinement
$H_{eff}^2\frac{\mathrm{d}P}{\mathrm{d}x}=(N_y-2)^2\frac{\mathrm{d}P}{\mathrm{d}
x } =Const$.
{\color{red} Michal, please add some results here and description to it as well.
The results for the force driven flows velocity phase profiles are shown in
Fig. \ref{fig:grid:profiles}. The corresponding ratios to the effective channel
width (without bounce-back nodes) is as $0.0809472165313,
0.0708111118485, 0.0678483852412, 0.0664499716904$. One can see
that for the channel widths as $149,198$, the results are convergent. The
velocities in the center of the bubble are $0.004176469553,
0.00401004133, 0.003969909065, 0.00394134433$, which corresponds
to the $Ca=\frac{\mu_{liq} U_{bubble}}{\gamma}=\frac{\frac{2}{3}\,
0.0045}{0.03771}=0.069$, which is slightly higher when for the assumed
profile. However, if we define Capillary number through the velocity of the
slug, then the Capillary number will have another value. The velocities of the
slug are $0.003411444831,0.003412693869,0.003420338754,0.003425757312$ and
Capillary number defined through these velocities is $0.0602$. 
}
Let us calculate to which extent we need to refine the interface to obtain
convergent results. The interface itself occupies approximately $5 \xi$, where
$\xi=\sqrt{k/A}=1$. This interface width divided to the effective channel width
($5\xi/H_{eff}$) is as follows {\color{red} Please add new values here
$0.102,0.051,0.034,0.025$ } for the given
$H_{eff}$. Let us calculate the interface width ratio to the film thickness as
$5\xi/H_{film}$ is as follows
{\color{red} Please add here new values $1.2605,  0.7205,  0.501,  0.3839$}.
Therefore, one needs to
resolve the interface as {\color{red} Please give number here $40\%$} of the
expected film width for simulations to be converged.
{\color{red} If we need this pictures???
\begin{figure}
\includegraphics[width=0.47\textwidth]{Figures/grid_phase_prof_49.eps}\hfill
\includegraphics[width=0.47\textwidth]{Figures/grid_phase_prof_98.eps}\\
\includegraphics[width=0.47\textwidth]{Figures/grid_phase_prof_147.eps}\hfill
\includegraphics[width=0.47\textwidth]{Figures/grid_phase_prof_196.eps}\\
\caption{Do we need these pictures? Grid refined profiles for the effictive
channel widths
$H_{eff}=49,98,147,196$.\label{fig:grid:profiles}}
\end{figure}
}

>>>>

The example what we want to simulate is the capillary number
$0.05$. As it was pointed out the film thickness in that case is $6$ percent.
Note that in the case of half-way bounce-back walls which are used in the
simulations one needs to calculate the film thickness as: 
\begin{equation}
\delta=(\phi_0-0.5)/(N_y-2),
\end{equation}
where $\phi_0$ is the grid coordinate where phase field equals to $0$, $N_y-2$
is related to the effective channel height. Note that it's a big
simplification to impose the boundary in the middle beetween the bounce-back
node and the fluid node which give rise to the viscosity dependent location of
the wall \cite{ginzburg-multireflection}. In the case of $\delta=0.06$ the
initial grid number
is $N_y=102$, which gives the horizontal grid number as $N_x-1=15(N_y-2)=1500$.
The bubble is initialized as a rectangular box with coordinates
$y=7\dots N_y-8$, $x=\frac{N_x}{3}\dots \frac{2 N_x}{3}$ and phase
$\phi_{bubble}=-1$. All other nodes are initialized with the phase field
$\phi=1$. The force gradient can be estimated through the Poiseuille
profile formula, Eq.\ref{poiseuille:velocity:center}, and it equals
$1.508\,10^{-6}$.

After choosing the reference parameters, the grid refinement needs to keep the
same the same macroscopic parameters as capillary number and viscosities ratio.
Therefore the simulations were performed with grid
numbers $N_y=102,127,152,177,202,227$, corresponding to the
effective channels widths as $H_{eff}=100,125,150,175,200,225$,
$N_x=1501,1876,2251,2626,3001,3376=15
H_{eff}+1$. It is easy to check
that keeping all liquid viscosities the same yields the following quantity to
be a
constant while doing the grid refinement
$H_{eff}^2\frac{\mathrm{d}P}{\mathrm{d}x}=(N_y-2)^2\frac{\mathrm{d}P}{\mathrm{d}
x } = Const$. We performed a number of simulations for the prescribed grids.
The unified scaled profiles are shown in Fig. \ref{fig:grid:profiles}. One can
see the visual results convergence for $H_{eff}\geq 175$. The calculated widths
are $0.0606$,$0.0618$,$0.0617$,$0.0629$,$0.0624$,$0.0624$. We accept the
convergent results for grids larger than $H_{eff}=175$. However, one can see
that imposing proper initialization technique, large enough time and the wall
wettability reveal results which are different only in 3rd digit even for
underresolved film widths. To calculate how good the interface thickness is
resolved the interface thickness to the film width thickness is calculated.The
interface itself occupies approximately $5 \xi$, where
$\xi=\sqrt{k/A}=1$. Let us calculate the interface width ratio to the film
thickness as
$5\xi/H_{film}$ is as follows $0.824$, $0.646$,  $0.539$, $0.453$,
$0.400$, $0.355$. Therefore, one needs to
resolve the interface as $40-50$ percent of the
expected film width for simulations to be converged. We further examine the
velocities in the center of the bubble to calculate the Capillary number. The
calculated numbers are $0.0041$, $0.0041$,$0.0040$,$0.0040$,$0.0041$,$0.0039$.
This corresponds to the average capillary number as $0.07$, which is slightly
larger than the expected capillary number. The corresponding error can be
attributed to the density and viscosity ratio which is much larger in the
Bretherton problem.
\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/Grid/norm_grid_profs.eps}
\caption{Grid refined profiles for the effictive
channel widths
$H_{eff}=100,125,150,175,200,225$.One can see
convergent results even for the underresolved film
thickness.\label{fig:grid:profiles}}
\end{figure}
